---
title: "masking_pca_mcr_als_confirmation"
author: "Anna Ulanova"
date: "2024-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Testing out the method described in Ruffing et al.

## R Markdown

```{r}
library(raster)

library(tidyverse)
library(ALS)
library(summarytools)
library(ggplot2)
library(userfriendlyscience)
library(car)
library(dplyr)
start = Sys.time()
```
1. Isolates all of the directories for VNIR and groups based on treatment, metal, grass
- control Se, control As, 5 ppm Se & As, 15 ppm Se & As, 25 ppm Se & As
-- Extract all directories that contain the file, raw_rd_rf
-- Iterates by directories
-- Separates by "/", and looks at the two elements
-- Looks up the reference sheet by finding the row that has matching GH ID an tray number
-- Isolates grass type, treatment, [ppm]
-- Assigns As/Se for first separated element
-- 
--EX: D:\VNIR\GH_20230724\16_3_20230724_2023_07_24_13_10_38

```{r Polished and debugged, check whether As or Sel}
grouping_VNIR_files <- function() {
  # CHANGE BACK
  # spectral_files <-
  #   list.files(
  #     path = 'C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands',
  #     pattern='subraster',
  #     recursive = TRUE,
  #     full.names = TRUE  
  #   )
  spectral_files <-
    list.files(
      path = 'D:/VNIR',
      pattern = 'raw_rd_rf$',
      recursive = TRUE,
      full.names = TRUE
    )
  # CHANGE BACK
  reference_dir_txt <-
    read.csv(
      'D:/reference_sheet.csv')
    # reference_dir_txt <-
    # read.csv(
    #   'C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands/reference_sheet - Copy.csv', sep=',')
  group_summary <- list('Dir', 'Group')
  # Creates a summary chart with each files directory and group ID
  for (file in spectral_files) {
    row <-
      reference_dir_txt[sapply(reference_dir_txt$Partial.directory, function(x)
        grepl(x, file, fixed = TRUE)),]
    appending <- list(file, row$GroupAssigned)
    group_summary = rbind(group_summary, appending)
  }
  # Groups by group ID
  dfgroup_summary <- as.data.frame(group_summary[-1, ])
  colnames(dfgroup_summary) <- group_summary[1, ]
  # grouped<-dfgroup_summary %>% group_by(Group) %>%
  # group_split()
  # dfgroup_summary<-reference_dir_txt[1:2,10:11]
  print('grouping_VNIR_files-- complete')
  assigning_control_v_treated(dfgroup_summary)
}
```


```{r Polished and debugged}
assigning_control_v_treated<-function(grouped){
  # Define the control groups and their corresponding treatments
  control_treatment_mapping <- list(
    SG_Sel_0 = c("SG_Sel_5", "SG_Sel_15", "SG_Sel_25"),
    bahia_Sel_0 = c("bahia_Sel_5", "bahia_Sel_15", "bahia_Sel_25"),
    PRG_Sel_0 = c("PRG_Sel_5", "PRG_Sel_15", "PRG_Sel_25"),
    SG_As_0 = c("SG_As_5", "SG_As_15", "SG_As_25"),
    bahia_As_0 = c("bahia_As_5", "bahia_As_15", "bahia_As_25"),
    # CHANGE BACK
    PRG_As_0 = c("PRG_As_5", "PRG_As_15", "PRG_As_25"))
    #PRG_As_0 = c("PRG_As_15"))

  control_type_names_mapping <- list(
    SG_Sel_0 = c("SG_Sel_control_v_5", "SG_Sel_control_v_15", "SG_Sel_control_v_25"),
    bahia_Sel_0 = c("bahia_Sel_control_v_5", "bahia_Sel_control_v_15", "bahia_Sel_control_v_25"),
    PRG_Sel_0 = c("PRG_Sel_control_v_5", "PRG_Sel_control_v_15", "PRG_Sel_control_v_25"),
    SG_As_0 = c("SG_As_control_v_5", "SG_As_control_v_15", "SG_As_control_v_25"),
    bahia_As_0 = c("bahia_As_control_v_5", "bahia_As_control_v_15", "bahia_As_control_v_25"),
    # CHANGE BACK
    PRG_As_0 = c("PRG_As_control_v_5", "PRG_As_control_v_15", "PRG_As_control_v_25"))
    #PRG_As_0 = c("PRG_As_control_v_15"))

  # Iterate over each control group in the mapping
  for (control in names(control_treatment_mapping)) {
    # Extract the control directories
    control_dirs <- grouped$Dir[grouped$Group == control]

    # Extract the treatment groups for this control
    treatment_groups <- control_treatment_mapping[[control]]

    # Loop through each treatment group and match correctly
    treatment_dirs_list <- list()

    for (i in seq_along(treatment_groups)) {
      treatment_group <- treatment_groups[i]
      treatment_dirs <- grouped$Dir[grouped$Group == treatment_group]

      # Add the treatment directories to the list
      treatment_dirs_list[[i]] <- treatment_dirs
    }

    # Ensure the directories are correctly ordered and mapped to respective treatments
    # CHANGE BACK
    treatment5 <- treatment_dirs_list[[1]]
    treatment15 <- treatment_dirs_list[[2]]
    treatment25 <- treatment_dirs_list[[3]]

    # Call the launching_analysis function with the correct treatment type mapping
    # CHANGE BACK
    launching_analysis(control_dirs, treatment5, control_type_names_mapping[[control]][1])
    launching_analysis(control_dirs, treatment15, control_type_names_mapping[[control]][2])
    launching_analysis(control_dirs, treatment25, control_type_names_mapping[[control]][3])
  }
}
```
- Takes the directories for each group, and completes pre-processing as well as analysis for control vs grouped
```{r}

launching_analysis <- function(group1, group2, type) {
    raster_group_summary <- matrix(1, nrow = 1, ncol = 59)
    # group= list of directories, raster_group_summary = blank summary matrix, type="control_vs_treatment_concentration"
    control <-
      open_combine_rasters(group1, raster_group_summary, 'control')
    treated <-
      open_combine_rasters(group2, raster_group_summary, 'treatment')
    print('Rasters have been combined')
    initial_comp <- complete_pca(control)
    #cat('Initial component is: ', initial_comp)
    print('PCA success')
    control_mcr <- complete_mcr_als(control, initial_comp, '')
    #cat('MCR component is: ', control_mcr$comp)
    print('Control MCR success')
    control_spectra <- control_mcr$spectra
    #write.csv(control_spectra,'D:/VNIR/control_spectral.csv')
    final_comp <- control_mcr$comp
    treated_mcr <- complete_mcr_als(treated, final_comp, '')
    print('Treated MCR success')
    treated_spectra <- treated_mcr$spectra
    #cat('Treated spectra: ', dim(treated_spectra), '\n')
    #cat('Control spectra: ', dim(control_spectra), '\n')
    output <-
      complete_howells_game(control_spectra, treated_spectra, type)
    
}
```

Iterates by each raster directory, applying mask and filter, converting to 2D, appending to overall group 
```{r}
open_combine_rasters <-
  function(summary_per_group,
           raster_group_summary,
           type) {
    # Creates empty matrix that will follow the converted 2D matrix form of (col*row, nbands)
    # Since the sampled matrix will be 100*100*59, we will create 1,10000
    for (dir in summary_per_group) {
      print(dir)
      raster <- stack(dir)
      mask <- creating_conditions(raster)
      print('mask created')
      mask_red <- mask$red
      mask_ob_avg <- mask$ob
      mask_r_bg_avg <- mask$rbg
      
      reclassified_raster <-
        classify_raster_based_on_conditions(mask_red, mask_ob_avg, mask_r_bg_avg)
      subset <- stacking_selected_bands(raster)
      masked_subset <-
        applying_mask_on_subset(subset, reclassified_raster)
      # cropped raster should have dimensions of [100,100,59]--> [10000,59]
      cropped_raster <- applying_spatial_subset(masked_subset)
      print(dim(cropped_raster))
      cropped_2D <- convert_to_2d(cropped_raster)
      #cat('Dimension of 2D cropped df: ',dim(cropped_2D),'\n')
      #cat('Dimension of raster group summary df:', dim(raster_group_summary),'\n')
      #print(raster_group_summary)
      print('open raster function--check point 1')
      raster_group_summary <- rbind(raster_group_summary, cropped_2D)
      print('open raster function--check point 2')
    }
    
    raster_group_summary <- raster_group_summary[-1, ]
    cat('Dimension of raster group summary df:',
        dim(raster_group_summary),
        '\n')
    write.csv(
      raster_group_summary,
      paste('D:/Reflectance summary/one_matrix', "_", type, ".txt")
    )
    return(raster_group_summary)
  }
```

Finds the 100*100 spatial subset of the raster by checking if the values are predominantly nonzero
```{r}

applying_spatial_subset <- function(raster) {
  large_matrix <- raster[[1]]
  nrow_large <- dim(raster)[1]
  ncol_large <- dim(raster)[2]
  # Size establishing the dimensions of the submatrix 100*100
  submatrix_size <- 100
  # Convert matrix to binary (1 for non-zero, 0 for zero) for easier counting
  binary_matrix <- large_matrix != 0
  
   if (submatrix_size > nrow_large || submatrix_size > ncol_large) {
    stop("Submatrix size is larger than the large matrix dimensions.")
  }
  
  # Create a cumulative sum matrix to store non-zero counts up to each element
  cumsum_matrix <- matrix(0, nrow = nrow_large + 1, ncol = ncol_large + 1)
  
  # Fill in the cumulative sum matrix
  for (i in 1:nrow_large) {
    for (j in 1:ncol_large) {
      cumsum_matrix[i + 1, j + 1] <- binary_matrix[i, j] + 
                                     cumsum_matrix[i, j + 1] + 
                                     cumsum_matrix[i + 1, j] - 
                                     cumsum_matrix[i, j]
    }
  }
  
  # Initialize variables to track the best submatrix
  max_nonzero_count <- -1
  best_row <- 1
  best_col <- 1
  
  # Loop through all possible top-left corners of the submatrix
  for (i in 1:(nrow_large - submatrix_size + 1)) {
    for (j in 1:(ncol_large - submatrix_size + 1)) {
      # Calculate the number of non-zero elements in the current submatrix
      total_nonzero <- cumsum_matrix[i + submatrix_size, j + submatrix_size] -
                       cumsum_matrix[i + submatrix_size, j] -
                       cumsum_matrix[i, j + submatrix_size] +
                       cumsum_matrix[i, j]
      
      # Update the best submatrix if this one has more non-zero elements
      if (total_nonzero > max_nonzero_count) {
        max_nonzero_count <- total_nonzero
        best_row <- i
        best_col <- j
      }
    }
  }
  
  # Check the final bounds before cropping
  if ((best_row + submatrix_size - 1) > nrow_large || 
      (best_col + submatrix_size - 1) > ncol_large) {
    stop("Submatrix coordinates exceed the dimensions of the large matrix.")
  }
  
  # Crop and return the submatrix with the most non-zero values
  cropped_raster <- raster::crop(raster, extent(raster,best_row,best_row + 99, 
                                    best_col,best_col+99))
  
  return(cropped_raster)
}
  # # cat('selected raster is: ' , dim(raster),'\n')
  # predominantly_zero = TRUE
  # col_range = seq(75,(columns-75),75)
  # row_range = seq(75,(rows-75),75)
  # coordinates<-expand.grid(x=row_range, y=col_range)
  # # subset_raster<-matrix(0, nrow=1, ncol=59)
  # # cat("parent raster is: ", dim(subset_raster),'\n')
  # while (predominantly_zero) {
  #   # Add no repeats
  #   shuffled_coordinates <- coordinates[sample(nrow(coordinates),1), ]
  #   random_r <- shuffled_coordinates$x
  #   random_c <- shuffled_coordinates$y
  #   row_min <- random_r - 49
  #   row_max <- random_r + 50
  #   col_min <- random_c - 49
  #   col_max <- random_c + 50
  #   # Extract the 100x100 square
  #   subset_raster <-
  #     crop(first_layer,
  #          extent(first_layer, row_min, row_max, col_min, col_max))
  #   
  #   # Convert the square to a matrix for easier analysis
  #   square_values <- as.matrix(subset_raster)
  #   
  #   # Count zero values
  #   zero_count <- sum(square_values == 0)
  #   # Count total number of cells in matrix
  #   total_count <-
  #     total_count <- dim(square_values)[1] * dim(square_values)[2]
  #   
  #   # Calculate the percentage of non-zero values
  #   zero_percentage <- (zero_count / total_count) * 100
  #   # Check if it's 60% or more
  #   if (zero_percentage <= 40) {
  #     cat(random_r,',', random_c,'\n')
  #     predominantly_zero = FALSE
  #     raster <-
  #       crop(raster,
  #            extent(raster, row_min, row_max, col_min, col_max))
  #     # Dimensions of cropped 3D should always be 100*100*59
  #     cat("DIMENSIONS OF 3D CROPPED RASTER: ", dim(raster), '\n')
  #     return(raster)
  #   } else{
  #       #print('')
  #   }
  # }
# }
```
2. For each raster per group: 
- Applies mask based on wavelength conditions
- Assigns random point and checks if (1) the 100 * 100 square exists on raster
(2) there are predominantly nonzero values in the outlined square 
- Completes spatial subset of the raster and isolates reflectance values for all bands
- Isolates the 59 bands used to make the masking conditions and creates a 100*100*59 raster
- Repeats the process for each file per group, stacking the subsetted rasters together
-- EX: if there are five files per group, the final raster will be 100*100*(59+5)




3.Uses the compiled rasters for control vs treated groups for the rest of the steps
- PCA
- MCR-ALS
- Games-Howell




FINDS AND AVERAGES BANDS OF INTEREST
FINDS REFLECTANCE RATIOS BASED ON WAVELENGTH CONDITIONS
COMPLETES OTSU'S METHOD
```{r}


creating_conditions <- function(Hyp_As_col3_16_3) {
  cat('creating conditions')
  #Averaging red between 750 and 850nm
  Red_avg <- mean(Hyp_As_col3_16_3[[180:204]])
  #Averaging orange between 550 and 600
  Orange_avg <- mean(Hyp_As_col3_16_3[[75:92]])
  #Averaging blue between 410 and 450nm
  Blue_avg <- mean(Hyp_As_col3_16_3[[10:25]])
  #Averaging for a blue-green band, in the paper it claimes between 3 and 450nm which makes no sense.
  #So I'm taking the liberty to make it between 450 and 550
  Blue_Green_avg <- mean(Hyp_As_col3_16_3[[30:69]])
  #Orange Average to Blue Average
  O_B_avg <- Orange_avg / Blue_avg

  #Red average to Blue-Green Average
  R_BG_avg <- Red_avg / Blue_Green_avg
  #Red Cutoff Threshold
  # Step 1: Calculate the histogram
  hist_values <-
    hist(getValues(Red_avg), breaks = 256, plot = FALSE)
  
  # Step 2: Implement Otsu's method
  # Compute probability distribution
  p <- hist_values$counts / sum(hist_values$counts)
  omega <- cumsum(p)
  mu <- cumsum(p * hist_values$mids)
  mu_t <- mu[length(mu)]
  
  # Compute between-class variance
  sigma_b_squared <- (mu_t * omega - mu) ^ 2 / (omega * (1 - omega))
  
  # Find the maximum value of between-class variance
  max_sigma_idx <- which.max(sigma_b_squared)
  otsu_thresh <- hist_values$mids[max_sigma_idx]
  
  # Step 3: Apply the threshold
  Red_cutoff <-
    calc(
      Red_avg,
      fun = function(x) {
        ifelse(x > otsu_thresh, 1, 0)
      }
    )
  # Plot to visualize
  # plot(Red_cutoff)
  return(list(red = Red_cutoff, ob = O_B_avg, rbg = R_BG_avg))
}

```


CREATES BINARY MASK BASED ON CONDITIONS
```{r}

classify_raster_based_on_conditions <-
  function(red_cutoff_raster,
           orange_blue_avg_raster,
           red_bg_avg_raster) {
    # Define a function to apply to each cell
    reclassify_function <-
      function(red_cutoff,
               orange_blue_avg,
               red_bg_avg) {
        # Check the conditions and assign 1 or 0
        condition1 <- red_cutoff == 1
        condition2 <- (orange_blue_avg >= 1.75) | (red_bg_avg > 2.5)
        
        result <- ifelse(condition1 & condition2, 1, 0)
        return(result)
      }
    
    # Stack the rasters for easier cell-wise operations
    raster_stack <-
      stack(red_cutoff_raster,
            orange_blue_avg_raster,
            red_bg_avg_raster)
    
    # Apply the reclassification function to each cell
    reclassified_raster <- calc(
      raster_stack,
      fun = function(x) {
        reclassify_function(x[1], x[2], x[3])
      }
    )
    # plot(reclassified_raster)
    return(reclassified_raster)
  }

```

STACKS ONLY RGB RASTERS
```{r}


stacking_selected_bands <- function(raster) {
  #range<-list([10:25],[30:69], [75:92],[180:204])
  red_stack <- raster[[180:204]]
  orange_stack <- raster[[75:92]]
  blue_green_stack <- raster[[30:69]]
  blue_stack <- raster[[10:25]]
  subset_raster <- stack(red_stack, orange_stack, blue_stack)
  print('stacking_selected_bands--complete')
  return(subset_raster)
}

```

APPLIES BINARY MASK ONTO BAND-SUBSETTED RASTER

```{r}

applying_mask_on_subset <- function(mask, raster) {
  masked <- mask * raster
  print('applying_mask_on_subset--complete')
  return(masked)
}
```


CONVERTS HYPERCUBE TO 2D MATRIX
```{r}

convert_to_2d <- function(hypercube) {
  print('lil cube')
  matrix2D <-
    matrix(
      hypercube,
      nrow = dim(hypercube)[1] * dim(hypercube)[2],
      ncol = dim(hypercube)[3]
    )
  print('convert_to_2d --complete')
  return(matrix2D)
}
```

COMPLETES PCA, FINDS EIGENVALUES, USES KAISERS RULE TO DETERMINE NUMBER OF COMPONENTS
```{r}

complete_pca <- function(matrix) {
  pca_result <- prcomp(matrix, scale. = TRUE)
  
  # Extract eigenvalues
  eigenvalues <- pca_result$sdev ^ 2
  # Remove comment if you want to see the Screen Plot
  # Create a scree plot to visualize eigenvalues and determine the number of components
  # plot(
  #   eigenvalues,
  #   type = "b",
  #   xlab = "Component",
  #   ylab = "Eigenvalue",
  #   main = "Scree Plot"
  # )
  # abline(h = 1, col = "red", lty = 2)  # Kaiser's rule threshold
  
  # Uses Kaiser-Guttman rule which suggest that "the elbow" in scree plots occurs
  # when the eigenvalues dip below 1.
  # https://datasciencewiki.net/kaisers-rule/
  num_components <- sum(eigenvalues > 1)
  return(num_components)
}
```

<!-- CONVERT ZEROS TO NANS -->
<!-- 1) CONVERTS ALL ZEROS IN 2D MATRIX INTO NANS -->
<!-- ```{R} -->
<!-- convert_to_zero<- function(matrix2D){ -->
<!--   nan_matrix<-gsub(0,NA, matrix2D) -->
<!--   return(nan_matrix) -->
<!-- } -->
<!-- ``` -->

CHUNKY MCR TASKS
1) CREATES INITIAL CONCENTRATION/SPECTRAL GUESSES
2) COMPLETES MCR-ALS
3) PLOTS ORIGINAL DATA VS RECONSTRUCTED SPECTRA
4) OUTPUTS NUMBER OF RELEVANT COMPONENTS FROM RECONSTRUCTED DATASET

```{r}

complete_mcr_als<-function(matrix2D, component, type){
  # Step 1: Perform MCR-ALS on control_2d
  # Initialize random matrices for MCR-ALS
  initial_concentration_profiles <-
    matrix(runif(nrow(matrix2D) * component),
           nrow = nrow(matrix2D),
           ncol = component)
  initial_spectra <- matrix(runif(component * ncol(matrix2D)),
                            nrow = ncol(matrix2D),
                            ncol = component)
  
  # Run MCR-ALS on control_2d
  mcr_result <- als(
    CList = list(initial_concentration_profiles),
    S = initial_spectra,
    PsiList = list(matrix2D),
    nonnegC = TRUE,
    nonnegS = TRUE,
    maxiter = 75,
    thresh = 1e-6
  )
  # Extract the resolved spectral profiles from mcr output
  resolved_spectra <- mcr_result$S
  resolved_concentration <- mcr_result$C[[1]]
  
  reconstructed_spectra <-
    resolved_concentration %*% t(resolved_spectra)
  #
  # Compares observed_spectra, input to the MCR-ALS algorithm [n_samples, n_bands], to reconstructed results
  # Example for plotting the original vs. reconstructed data
  n_samples <- nrow(matrix2D)
  n_bands <- ncol(matrix2D)
  # Remove comment if you want to plot the MCR output
  # title <-
  #   paste0(type, ": Comparison of Original and Reconstructed Spectra")
  # save_name <-
  #   paste0(
  #     'C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/',
  #     type,
  #     '_mcr_als.jpg'
  #   )
  # Plot for the first sample
  # jpeg(file = save_name)
  # plot(
  #   1:n_bands,
  #   matrix2D[3,],
  #   type = "l",
  #   col = "red",
  #   xlab = "Wavelength/Band",
  #   ylab = "Intensity",
  #   main = title
  # )
  # lines(1:n_bands, reconstructed_spectra[3,], col = "blue")
  # legend(
  #   "topright",
  #   legend = c("Original", "Reconstructed"),
  #   col = c("red", "blue")
  # )
  # dev.off()
  #Complete PCA to find if the number of components has changed
  # # Eigenvalue calculation (using concentration profiles as an example)
  eigenvalues <- eigen(cov(reconstructed_spectra))$values
  
  # Kaiser’s rule: Select components with eigenvalues > 1
  significant_components <- which(eigenvalues > 1)
  return(list(comp = significant_components, spectra = resolved_spectra))
}
```

GAMES-HOWELL
```{r}
complete_howells_game<-function(control, treated, name){
  cat('-----Games-Howell Test Debugging-----', '\n')
  final_spectra_avg <- rowMeans(control)  # This takes the average of the two columns in final_spectra
  
  # Ensure that both final_spectra_avg and treated_spectra are vectors of the same length
  # Combine them into a single dataframe with a group label
  
  control_spectra <- data.frame(Spectra = final_spectra_avg, Group = "Control")
  treated_spectra_df <- data.frame(Spectra = treated, Group = "Treated")
  
  # Combine the two datasets into one dataframe
  compiled_spectra <- rbind(control_spectra, treated_spectra_df)
  # write.csv(compiled_spectra, 'D:/Reflectance summary/compiled_spectra.csv')
  # Perform the Games-Howell test
  gh_results <- posthocTGH(y = compiled_spectra$Spectra, 
                           x = compiled_spectra$Group, 
                           method = "games-howell")
  
  # Extract mean and confidence intervals for plotting
  means <- tapply(compiled_spectra$Spectra, compiled_spectra$Group, mean)
  ci <- tapply(compiled_spectra$Spectra, compiled_spectra$Group, function(x) {
    c(
      mean(x) - qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x)),
      mean(x) + qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x))
    )
  })
  # 
  # # Prepare data for plotting
  # plot_data <- data.frame(
  #   Group = names(means),
  #   Mean = as.numeric(means),
  #   CI_Lower = sapply(ci, `[`, 1),
  #   CI_Upper = sapply(ci, `[`, 2)
  # )
  # 
  # # Create the plot
  # ggplot(plot_data, aes(x = Group, y = Mean)) +
  #   geom_point(size = 3) +
  #   geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
  #   labs(title = "Mean Spectral Values by Group with 95% Confidence Intervals",
  #        x = "Group", y = "Mean Spectral Value") +
  #   theme_minimal()
  # 
  
  # 4. Prepare a summary dataframe
  summary_df <- data.frame(
    Control = means[1],
    Treated = means[2],
    Diff = gh_results$output$games.howell$diff,
    CI_Lower = gh_results$output$games.howell$ci.lo,
    CI_Upper = gh_results$output$games.howell$ci.hi,
    P_Value = gh_results$output$games.howell$p
  )
  
  # 5. Save the summary to a CSV file
  write.csv(summary_df, paste('D:/Reflectance summary/', name, '_summary.txt'), row.names = FALSE)
}
```
MAIN
```{r}

# control<-stack('C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands/14_1_20230724_2023_07_24_12_33_04/raw_rd_rf')
# treated<-stack('C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands/4_1_20230724_2023_07_24_09_00_14/raw_rd_rf')
# 
# control_mask<-creating_conditions(control)
# control_red_cutoff<-control_mask$red
# control_ob_avg<-control_mask$ob
# control_r_bg_avg<-control_mask$rbg
# 
# control_reclassified_raster <- classify_raster_based_on_conditions(red_cutoff_raster = control_red_cutoff, orange_blue_avg_raster = control_ob_avg, red_bg_avg_raster = control_r_bg_avg)
# plot(control_reclassified_raster)
# #
# treated_mask<-creating_conditions(treated)
# treated_red_cutoff<-treated_mask$red
# treated_ob_avg<-treated_mask$ob
# treated_r_bg_avg<-treated_mask$rbg
# 
# treated_reclassified_raster <- classify_raster_based_on_conditions(red_cutoff_raster = treated_red_cutoff, orange_blue_avg_raster = treated_ob_avg, red_bg_avg_raster = treated_r_bg_avg)
# 
# 
# plot(treated_reclassified_raster)
# 
# 
# control_subset<-stacking_selected_bands(control)
# 
# treated_subset<-stacking_selected_bands(treated)
# 
# masked_control_subset<-applying_mask_on_subset(control_subset, control_reclassified_raster)
# masked_treated_subset<-applying_mask_on_subset(treated_subset, treated_reclassified_raster)

# subset_control<-applying_spatial_subset(masked_control_subset)
# 
# control2d<-convert_to_2d(masked_control_subset)
# treated2d<-convert_to_2d(masked_treated_subset)
# 
# write.csv(control2d, 'C:\\Users\\RDCRLAAU\\Desktop\\Plant_as_bioindicators\\MCR-ALS R Processing\\subset_testing_control.txt')
# write.csv(treated2d, 'C:\\Users\\RDCRLAAU\\Desktop\\Plant_as_bioindicators\\MCR-ALS R Processing\\subset_testing_treatment.txt')
# initial_comp<-complete_pca(control2d)
# #
# # na_control<-convert_to_zero(control2d)
# #
# # na_treated<- convert_to_zero(treated2d)
# 
# control_mcr<-complete_mcr_als(control2d, initial_comp,'Control')
# control_spectra<-control_mcr$spectra
# # control spectra: [1:59, 1:2]
# treated_mcr<-complete_mcr_als(treated2d, control_mcr$comp, 'Treated')
# treated_spectra<-treated_mcr$spectra
# treated_spectra<-treated_mcr$spectra
# # treated spectra: [1:59,1]
# print(Sys.time()-start) # Takes 33.9 min to complete
# output<-complete_howells_game(control_spectra, treated_spectra)
# plotGamesHowell
grouping_VNIR_files()
print(Sys.time()-start)
```