---
title: "masking_pca_mcr_als_confirmation"
author: "Anna Ulanova"
date: "2024-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Testing out the method described in Ruffing et al.

## R Markdown

```{r}
library(raster)
library(tidyverse)
library(ALS)
library(summarytools)
library(ggplot2)
library(userfriendlyscience)
library(car)
library(dplyr)
start = Sys.time()
```
1. Isolates all of the directories for VNIR and groups based on treatment, metal, grass
- control Se, control As, 5 ppm Se & As, 15 ppm Se & As, 25 ppm Se & As
-- Extract all directories that contain the file, raw_rd_rf
-- Iterates by directories
-- Separates by "/", and looks at the two elements
-- Looks up the reference sheet by finding the row that has matching GH ID an tray number
-- Isolates grass type, treatment, [ppm]
-- Assigns As/Se for first separated element
-- 
--EX: D:\VNIR\GH_20230724\16_3_20230724_2023_07_24_13_10_38

```{r Polished and debugged}
grouping_VNIR_files<-function(){
  spectral_files<-list.files(path='D:/VNIR/GH_20231213', pattern='raw_rd_rf$', recursive=TRUE, full.names= TRUE)
  reference_dir_txt<- read.csv('C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/reference_sheet.txt', sep='\t')
  group_summary<-list('Dir','Group')
  # Creates a summary chart with each files directory and group ID
  for (file in spectral_files){
    row<-reference_dir_txt[sapply(reference_dir_txt$Partial.directory, function(x) grepl(x, file, fixed = TRUE)), ]
    print(row)
    appending<-list(file, row$GroupAssigned)
    group_summary=rbind(group_summary, appending)
  }
  # Groups by group ID
  dfgroup_summary<-as.data.frame(group_summary[-1,])
  colnames(dfgroup_summary)<-group_summary[1,]
  # grouped<-dfgroup_summary %>% group_by(Group) %>%
  # group_split()
  assigning_control_v_treated(dfgroup_summary)
}
```


```{r Polished and debugged}
assigning_control_v_treated<-function(grouped){
  
  #-----------------------As treatment-----------------------#
  # SG As run
  control<- grouped%>%filter(Group=='SG_As_0')
  control_2D<-launch_rasters(control)
  
  treatment5<-grouped%>%filter(Group=='SG_As_5')
  treatment5_2D<-launch_rasters(treatment5)
  type<-"SG_As_control_v_5ppm"
  
  treatment15<-grouped%>%filter(Group=='SG_As_15')
  treatment15_2D<-launch_rasters(treatment15)
  type<-"SG_As_control_v_15ppm"
  
  treatment25<-grouped%>%fiter(Group=='SG_As_25')
  treatment25_2D<-launch_rasters(treatment25)
  type<-"SG_As_control_v_25ppm"
  
  # PRG As run
  control<- grouped%>%filter(Group=='PRG_As_0')
  treatment5<-grouped%>%filter(Group=='PRG_As_5')
  type<-"PGR_As_control_v_5ppm"
  
  treatment15<-grouped%>%filter(Group=='PRG_As_15')
  type<-"PGR_As_control_vs_15ppm"
  
  treatment25<-grouped%>%fiter(Group=='PRG_As_25')
  type<-"PGR_As_control_v_25ppm"
  # Bahia As run
  control<- grouped%>%filter(Group=='bahia_As_0')
  control_2D<-launch_rasters(control)
  
  treatment5<-grouped%>%filter(Group=='bahia_As_5')
  type<-"bahia_As_control_v_5ppm"
  
  treatment15<-grouped%>%filter(Group=='bahia_As_15')
  type<-"bahia_As_control_v_15ppm"
  
  treatment25<-grouped%>%fiter(Group=='bahia_As_25')
  type<-"bahia_As_control_v_25ppm"
  #-----------------------Se treatment-----------------------#
  # SG Sel run
  control<- grouped%>%filter(Group=='SG_Sel_0')
  control_2D<-launch_rasters(control)
  treatment5<-grouped%>%filter(Group=='SG_Sel_5')
  type<-"SG_Sel_control_v_5ppm"
  
  treatment5_2D<-launch_rasters(treatment5)
  type<-"SG_Sel_control_5ppm"
  
  treatment15<-grouped%>%filter(Group=='SG_Sel_15')
  treatment15_2D<-launch_rasters(treatment15)
  type<-"SG_Sel_control_v_15pp"
  
  treatment25<-grouped%>%fiter(Group=='SG_Sel_25')
  treatment25_2D<-launch_rasters(treatment25)
  type<-"SG_Sel_control_v_25ppm"
  # PRG Sel run
  control<- grouped%>%filter(Group=='PRG_Sel_0')
  treatment5<-grouped%>%filter(Group=='PRG_Sel_5')
  type<-"PRG_Sel_control_v_5ppm"
  treatment15<-grouped%>%filter(Group=='PRG_Sel_15')
  type<-"PRG_Sel_control_v_15ppm"
  treatment25<-grouped%>%fiter(Group=='PRG_Sel_25')
  type<-"PRG_Sel_control_v_25ppm"
  # Bahia Sel run
  control<- grouped%>%filter(Group=='bahia_Sel_0')
  treatment5<-grouped%>%filter(Group=='bahia_Sel_5')
  type<-"bahia_Sel_control_v_5ppm"
  treatment15<-grouped%>%filter(Group=='bahia_Sel_15')
  type<-"bahia_Sel_control_v_15ppm"
  treatment25<-grouped%>%fiter(Group=='bahia_Sel_25') 
  type<-"bahia_Sel_control_v_ppm"
}
```

```{r}
launching_analysis<-function(group1, group2, type){
  control<-open_combine_rasters(group1)
  treated<-open_combine_rasters(group2)
  intial_comp<-complete_pca(control)
  control_mcr<-complete_mcr_als(control2d, initial_comp, '')
  final_comp<-control_mcr$spectra
  treated_mcr<-complete_mcr_als(treated, final_comp, '')
  treated_spectra<-treated_mcr$spectra
  treated_comp<-treated_mcr$comp
  output<-complete_howells_game(final_comp, treated_comp, type)
}
```

Iterates by each raster directory, applying mask and filter, converting to 2D, appending to overall group 
```{r}
open_combine_rasters<-function(summary_per_group){
  group_dirs<-summary_per_group$Dir
  # Creates empty matrix that will follow the converted 2D matrix form of (col*row, nbands)
  # Since the sampled matrix will be 100*100*59, we will create 1,10000
  raster_group_summary<-matrix(nrow=1, ncol=100*100)
  for (dir in group_dirs){
    raster<-stack(dir)
    mask<-creating_conditions(raster)
    mask_red<-mask$red
    mask_ob_avg<-mask$ob
    mask_r_bg_avg<-mask$rbg
    
    reclassified_raster<-classifying_raster_based_on_conditions(mask_red, mask_ob_avg, mask_r_bg_avg)
    subset<-stacking_selected_bands(raster)
    masked_subset<-applying_mask_on_subset(raster, reclassified_raster)
    cropped_raster<-applying_spatial_subset(masked_subset)
    cropped_2D<-convert_to_2d(cropped_raster)
    raster_group_summary<-rbind(raster_group_summary, cropped_2D)
  }
  return(raster_group_summary)
}
```

Finds the 100*100 spatial subset of the raster by checking if the values are predominantly nonzero
```{r}
applying_spatial_subset<-function(raster){
  first_layer<-raster[[1]]
  rows<-dim(raster)[1]
  columns<-dim(raster)[2]
  predominantly_zero=TRUE
  subset_raster<-matrix(0, nrow=100, ncol=100)
  while (predominantly_nonzero){
    random_r<-sample(100:(rows-100),1)
    random_c<-sample(100:(columns-100),1)
    row_min <-random_r-50
    row_max <-random_r+50
    col_min <-random_c-50
    col_max <-random_c+50
    # Extract the 100x100 square
    subset_raster <- crop(first_layer, extent(first_layer, col_min, col_max, row_min, row_max))
    
    # Convert the square to a matrix for easier analysis
    square_values <- as.matrix(subset_raster)
    
    # Count non-zero values
    non_zero_count <- sum(square_values != 0)
    total_count <- length(square_values)
    
    # Calculate the percentage of non-zero values
    non_zero_percentage <- (non_zero_count / total_count) * 100
    # Check if it's 60% or more
    if (non_zero_percentage >= 60){
      predominantly_zero=FALSE
      raster<-crop(raster, extent(raster, col_min, col_max, row_min, row_max))
      return(raster)
    }
  }
  
}
```
2. For each raster per group: 
- Applies mask based on wavelength conditions
- Assigns random point and checks if (1) the 100 * 100 square exists on raster
(2) there are predominantly nonzero values in the outlined square 
- Completes spatial subset of the raster and isolates reflectance values for all bands
- Isolates the 59 bands used to make the masking conditions and creates a 100*100*59 raster
- Repeats the process for each file per group, stacking the subsetted rasters together
-- EX: if there are five files per group, the final raster will be 100*100*(59+5)




3.Uses the compiled rasters for control vs treated groups for the rest of the steps
- PCA
- MCR-ALS
- Games-Howell




FINDS AND AVERAGES BANDS OF INTEREST
FINDS REFLECTANCE RATIOS BASED ON WAVELENGTH CONDITIONS
COMPLETES OTSU'S METHOD
```{r}

creating_conditions <- function(Hyp_As_col3_16_3) {
  #Averaging red between 750 and 850nm
  Red_avg <- raster::mean(Hyp_As_col3_16_3[[180:204]])
  
  #Averaging orange between 550 and 600
  Orange_avg <- raster::mean(Hyp_As_col3_16_3[[75:92]])
  
  #Averaging blue between 410 and 450nm
  Blue_avg <- raster::mean(Hyp_As_col3_16_3[[10:25]])
  
  #Averaging for a blue-green band, in the paper it claimes between 3 and 450nm which makes no sense.
  #So I'm taking the liberty to make it between 450 and 550
  Blue_Green_avg <- raster::mean(Hyp_As_col3_16_3[[30:69]])
  
  #Orange Average to Blue Average
  O_B_avg <- Orange_avg / Blue_avg
  
  #Red average to Blue-Green Average
  R_BG_avg <- Red_avg / Blue_Green_avg
  
  #Red Cutoff Threshold
  # Step 1: Calculate the histogram
  hist_values <- hist(getValues(Red_avg), breaks = 256, plot = FALSE)
  
  # Step 2: Implement Otsu's method
  # Compute probability distribution
  p <- hist_values$counts / sum(hist_values$counts)
  omega <- cumsum(p)
  mu <- cumsum(p * hist_values$mids)
  mu_t <- mu[length(mu)]
  
  # Compute between-class variance
  sigma_b_squared <- (mu_t * omega - mu) ^ 2 / (omega * (1 - omega))
  
  # Find the maximum value of between-class variance
  max_sigma_idx <- which.max(sigma_b_squared)
  otsu_thresh <- hist_values$mids[max_sigma_idx]
  
  # Step 3: Apply the threshold
  Red_cutoff <-
    calc(
      Red_avg,
      fun = function(x) {
        ifelse(x > otsu_thresh, 1, 0)
      }
    )
  # Plot to visualize
  plot(Red_cutoff)
  return(list(red = Red_cutoff, ob = O_B_avg, rbg = R_BG_avg))
}


```


CREATES BINARY MASK BASED ON CONDITIONS
```{r}

classify_raster_based_on_conditions <-
  function(red_cutoff_raster,
           orange_blue_avg_raster,
           red_bg_avg_raster) {
    # Define a function to apply to each cell
    reclassify_function <-
      function(red_cutoff,
               orange_blue_avg,
               red_bg_avg) {
        # Check the conditions and assign 1 or 0
        condition1 <- red_cutoff == 1
        condition2 <- (orange_blue_avg >= 1.75) | (red_bg_avg > 2.5)
        
        result <- ifelse(condition1 & condition2, 1, 0)
        return(result)
      }
    
    # Stack the rasters for easier cell-wise operations
    raster_stack <-
      stack(red_cutoff_raster,
            orange_blue_avg_raster,
            red_bg_avg_raster)
    
    # Apply the reclassification function to each cell
    reclassified_raster <- calc(
      raster_stack,
      fun = function(x) {
        reclassify_function(x[1], x[2], x[3])
      }
    )
    print(dim(reclassified_raster))
    return(reclassified_raster)
  }

```

STACKS ONLY RGB RASTERS
```{r}

stacking_selected_bands <- function(raster) {
  #range<-list([10:25],[30:69], [75:92],[180:204])
  red_stack <- raster[[180:204]]
  orange_stack <- raster[[75:92]]
  blue_green_stack <- raster[[30:69]]
  blue_stack <- raster[[10:25]]
  subset_raster <- stack(red_stack, orange_stack, blue_stack)
  return(subset_raster)
}

```

APPLIES BINARY MASK ONTO BAND-SUBSETTED RASTER

```{r}

applying_mask_on_subset <- function(mask, raster) {
  masked <- mask * raster
  return(masked)
}
```


CONVERTS HYPERCUBE TO 2D MATRIX
```{r}

convert_to_2d <- function(hypercube) {
  matrix2D <-
    matrix(
      hypercube,
      nrow = dim(hypercube)[1] * dim(hypercube)[2],
      ncol = dim(hypercube)[3]
    )
  return(matrix2D)
}
```

COMPLETES PCA, FINDS EIGENVALUES, USES KAISERS RULE TO DETERMINE NUMBER OF COMPONENTS
```{r}

complete_pca <- function(matrix) {
  pca_result <- prcomp(matrix, scale. = TRUE)
  
  # Extract eigenvalues
  eigenvalues <- pca_result$sdev ^ 2
  # Remove comment if you want to see the Screen Plot
  # Create a scree plot to visualize eigenvalues and determine the number of components
  # plot(
  #   eigenvalues,
  #   type = "b",
  #   xlab = "Component",
  #   ylab = "Eigenvalue",
  #   main = "Scree Plot"
  # )
  abline(h = 1, col = "red", lty = 2)  # Kaiser's rule threshold
  
  # Uses Kaiser-Guttman rule which suggest that "the elbow" in scree plots occurs
  # when the eigenvalues dip below 1.
  # https://datasciencewiki.net/kaisers-rule/
  num_components <- sum(eigenvalues > 1)
  return(num_components)
}
```

<!-- CONVERT ZEROS TO NANS -->
<!-- 1) CONVERTS ALL ZEROS IN 2D MATRIX INTO NANS -->
<!-- ```{R} -->
<!-- convert_to_zero<- function(matrix2D){ -->
<!--   nan_matrix<-gsub(0,NA, matrix2D) -->
<!--   return(nan_matrix) -->
<!-- } -->
<!-- ``` -->

CHUNKY MCR TASKS
1) CREATES INITIAL CONCENTRATION/SPECTRAL GUESSES
2) COMPLETES MCR-ALS
3) PLOTS ORIGINAL DATA VS RECONSTRUCTED SPECTRA
4) OUTPUTS NUMBER OF RELEVANT COMPONENTS FROM RECONSTRUCTED DATASET

```{r}

complete_mcr_als<-function(matrix2D, component, type){
  # Step 1: Perform MCR-ALS on control_2d
  # Initialize random matrices for MCR-ALS
  initial_concentration_profiles <-
    matrix(runif(nrow(matrix2D) * component),
           nrow = nrow(matrix2D),
           ncol = component)
  initial_spectra <- matrix(runif(component * ncol(matrix2D)),
                            nrow = ncol(matrix2D),
                            ncol = component)
  
  # Run MCR-ALS on control_2d
  mcr_result <- als(
    CList = list(initial_concentration_profiles),
    S = initial_spectra,
    PsiList = list(matrix2D),
    nonnegC = TRUE,
    nonnegS = TRUE,
    maxiter = 75,
    thresh = 1e-6
  )
  # Extract the resolved spectral profiles from mcr output
  resolved_spectra <- mcr_result$S
  resolved_concentration <- mcr_result$C[[1]]
  
  reconstructed_spectra <-
    resolved_concentration %*% t(resolved_spectra)
  #
  # Compares observed_spectra, input to the MCR-ALS algorithm [n_samples, n_bands], to reconstructed results
  # Example for plotting the original vs. reconstructed data
  n_samples <- nrow(matrix2D)
  n_bands <- ncol(matrix2D)
  # Remove comment if you want to plot the MCR output
  # title <-
  #   paste0(type, ": Comparison of Original and Reconstructed Spectra")
  # save_name <-
  #   paste0(
  #     'C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/',
  #     type,
  #     '_mcr_als.jpg'
  #   )
  # # Plot for the first sample
  # jpeg(file = save_name)
  # plot(
  #   1:n_bands,
  #   matrix2D[3,],
  #   type = "l",
  #   col = "red",
  #   xlab = "Wavelength/Band",
  #   ylab = "Intensity",
  #   main = title
  # )
  # lines(1:n_bands, reconstructed_spectra[3,], col = "blue")
  # legend(
  #   "topright",
  #   legend = c("Original", "Reconstructed"),
  #   col = c("red", "blue")
  # )
  # dev.off()
  #Complete PCA to find if the number of components has changed
  # # Eigenvalue calculation (using concentration profiles as an example)
  eigenvalues <- eigen(cov(reconstructed_spectra))$values
  
  # Kaiser’s rule: Select components with eigenvalues > 1
  significant_components <- which(eigenvalues > 1)
  return(list(comp = significant_components, spectra = resolved_spectra))
}
```

GAMES-HOWELL
```{r}
complete_howells_game<-function(control, treated, name){
  final_spectra_avg <- rowMeans(control)  # This takes the average of the two columns in final_spectra
  
  # Ensure that both final_spectra_avg and treated_spectra are vectors of the same length
  # Combine them into a single dataframe with a group label
  
  control_spectra <- data.frame(Spectra = final_spectra_avg, Group = "Control")
  treated_spectra_df <- data.frame(Spectra = treated, Group = "Treated")
  
  # Combine the two datasets into one dataframe
  compiled_spectra <- rbind(control_spectra, treated_spectra_df)
  
  # Perform the Games-Howell test
  gh_results <- posthocTGH(y = compiled_spectra$Spectra, 
                         x = compiled_spectra$Group, 
                         method = "games-howell")
  
  # Extract mean and confidence intervals for plotting
  means <- tapply(compiled_spectra$Spectra, compiled_spectra$Group, mean)
  ci <- tapply(compiled_spectra$Spectra, compiled_spectra$Group, function(x) {
    c(
      mean(x) - qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x)),
      mean(x) + qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x))
    )
  })
  
  # Prepare data for plotting
  plot_data <- data.frame(
    Group = names(means),
    Mean = as.numeric(means),
    CI_Lower = sapply(ci, `[`, 1),
    CI_Upper = sapply(ci, `[`, 2)
  )
  
  # Create the plot
  ggplot(plot_data, aes(x = Group, y = Mean)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
    labs(title = "Mean Spectral Values by Group with 95% Confidence Intervals",
         x = "Group", y = "Mean Spectral Value") +
    theme_minimal()
  
  
  # 4. Prepare a summary dataframe
  summary_df <- data.frame(
    Group1 = rownames(gh_summary),
    Control = group_means[1,2],
    Treated = group_means[2,2],
    Diff = gh_summary$diff,
    CI_Lower = gh_summary$ci.lo,
    CI_Upper = gh_summary$ci.hi,
    P_Value = gh_summary$p
  )
  
  # 5. Save the summary to a CSV file
  write.csv(summary_df, paste('D:/Reflectance summary/', name, '_summary.txt'), row.names = FALSE)
  
}
# complete_howells_game <- function(control, treated) {
#   # Combine the data into a single data frame
#   compiled_data <-
#     data.frame(
#       Group = rep(c("Treated", "Control1", "Control2"), each = 59),
#       Value = c(treated_spectra, control_spectra[, 1], control_spectra[, 2])
#     )
#   
#   # Perform Games-Howell test
#   #https://rdocumentation.org/packages/userfriendlyscience/versions/0.7.2/topics/posthocTGH
#   gh_test <-
#     posthocTGH(y = compiled_data$Value,
#                x = compiled_data$Group,
#                method = "games-howell")
#   return(list(test=gh_test, compiled_data=compiled_data))
# }
# ```
# 
# ```{r Optional chunk- plots only means}
# 
# plotGamesHowell <- function(compiled_data) {
#   # Extract mean and confidence intervals from the test results
#   means <- tapply(compiled_data$Value, compiled_data$Group, mean)
#   ci <- tapply(compiled_data$Value, compiled_data$Group, function(x) {
#     c(
#       mean(x) - qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x)),
#       mean(x) + qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x))
#     )
#   })
#   
#   # Prepare data for plotting
#   plot_data <- data.frame(
#     Group = names(means),
#     Mean = as.numeric(means),
#     CI_Lower = sapply(ci, `[`, 1),
#     CI_Upper = sapply(ci, `[`, 2)
#   )
#   
#   # Create the plot
#   ggplot(plot_data, aes(x = Group, y = Mean)) +
#     geom_point(size = 3) +
#     geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
#     labs(title = "Mean Spectral Values by Group with 95% Confidence Intervals",
#          x = "Group", y = "Mean Spectral Value") +
#     theme_minimal()
# }
```
MAIN
```{r}

control<-stack('C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands/14_1_20230724_2023_07_24_12_33_04/raw_rd_rf')
treated<-stack('C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands/4_1_20230724_2023_07_24_09_00_14/raw_rd_rf')

control_mask<-creating_conditions(control)
control_red_cutoff<-control_mask$red
control_ob_avg<-control_mask$ob
control_r_bg_avg<-control_mask$rbg

control_reclassified_raster <- classify_raster_based_on_conditions(red_cutoff_raster = control_red_cutoff, orange_blue_avg_raster = control_ob_avg, red_bg_avg_raster = control_r_bg_avg)


treated_mask<-creating_conditions(treated)
treated_red_cutoff<-treated_mask$red
treated_ob_avg<-treated_mask$ob
treated_r_bg_avg<-treated_mask$rbg

treated_reclassified_raster <- classify_raster_based_on_conditions(red_cutoff_raster = treated_red_cutoff, orange_blue_avg_raster = treated_ob_avg, red_bg_avg_raster = treated_r_bg_avg)


plot(treated_reclassified_raster)


control_subset<-stacking_selected_bands(control)

treated_subset<-stacking_selected_bands(treated)

masked_control_subset<-applying_mask_on_subset(control_subset, control_reclassified_raster)
masked_treated_subset<-applying_mask_on_subset(treated_subset, treated_reclassified_raster)

control2d<-convert_to_2d(masked_control_subset)
treated2d<-convert_to_2d(masked_treated_subset)

initial_comp<-complete_pca(control2d)
# 
# na_control<-convert_to_zero(control2d)
# 
# na_treated<- convert_to_zero(treated2d)

control_mcr<-complete_mcr_als(control2d, initial_comp,'Control')
final_spectra<-control_mcr$spectra
control_spectra<-control_mcr$spectra
# control spectra: [1:59, 1:2]
treated_mcr<-complete_mcr_als(treated2d, final_comp, 'Treated')
treated_spectra<-treated_mcr$spectra
treated_spectra<-treated_mcr$spectra
# treated spectra: [1:59,1]
print(Sys.time()-start) # Takes 33.9 min to complete
output<-complete_howells_game(final_spectra, treated_spectra)
# plotGamesHowell
```
