---
title: "masking_pca_mcr_als_confirmation"
author: "Anna Ulanova"
date: "2024-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Testing out the method described in Ruffing et al.

## R Markdown

```{r}
library(raster)
library(tidyverse)
library(ALS)
library(summarytools)
library(ggplot2)
library(userfriendlyscience)
library(car)
start = Sys.time()
```

FINDS AND AVERAGES BANDS OF INTEREST
FINDS REFLECTANCE RATIOS BASED ON WAVELENGTH CONDITIONS
COMPLETES OTSU'S METHOD
```{r}

creating_conditions <- function(Hyp_As_col3_16_3) {
  #Averaging red between 750 and 850nm
  Red_avg <- raster::mean(Hyp_As_col3_16_3[[180:204]])
  
  #Averaging orange between 550 and 600
  Orange_avg <- raster::mean(Hyp_As_col3_16_3[[75:92]])
  
  #Averaging blue between 410 and 450nm
  Blue_avg <- raster::mean(Hyp_As_col3_16_3[[10:25]])
  
  #Averaging for a blue-green band, in the paper it claimes between 3 and 450nm which makes no sense.
  #So I'm taking the liberty to make it between 450 and 550
  Blue_Green_avg <- raster::mean(Hyp_As_col3_16_3[[30:69]])
  
  #Orange Average to Blue Average
  O_B_avg <- Orange_avg / Blue_avg
  
  #Red average to Blue-Green Average
  R_BG_avg <- Red_avg / Blue_Green_avg
  
  #Red Cutoff Threshold
  # Step 1: Calculate the histogram
  hist_values <- hist(getValues(Red_avg), breaks = 256, plot = FALSE)
  
  # Step 2: Implement Otsu's method
  # Compute probability distribution
  p <- hist_values$counts / sum(hist_values$counts)
  omega <- cumsum(p)
  mu <- cumsum(p * hist_values$mids)
  mu_t <- mu[length(mu)]
  
  # Compute between-class variance
  sigma_b_squared <- (mu_t * omega - mu) ^ 2 / (omega * (1 - omega))
  
  # Find the maximum value of between-class variance
  max_sigma_idx <- which.max(sigma_b_squared)
  otsu_thresh <- hist_values$mids[max_sigma_idx]
  
  # Step 3: Apply the threshold
  Red_cutoff <-
    calc(
      Red_avg,
      fun = function(x) {
        ifelse(x > otsu_thresh, 1, 0)
      }
    )
  # Plot to visualize
  plot(Red_cutoff)
  return(list(red = Red_cutoff, ob = O_B_avg, rbg = R_BG_avg))
}


```


CREATES BINARY MASK BASED ON CONDITIONS
```{r}

classify_raster_based_on_conditions <-
  function(red_cutoff_raster,
           orange_blue_avg_raster,
           red_bg_avg_raster) {
    # Define a function to apply to each cell
    reclassify_function <-
      function(red_cutoff,
               orange_blue_avg,
               red_bg_avg) {
        # Check the conditions and assign 1 or 0
        condition1 <- red_cutoff == 1
        condition2 <- (orange_blue_avg >= 1.75) | (red_bg_avg > 2.5)
        
        result <- ifelse(condition1 & condition2, 1, 0)
        return(result)
      }
    
    # Stack the rasters for easier cell-wise operations
    raster_stack <-
      stack(red_cutoff_raster,
            orange_blue_avg_raster,
            red_bg_avg_raster)
    
    # Apply the reclassification function to each cell
    reclassified_raster <- calc(
      raster_stack,
      fun = function(x) {
        reclassify_function(x[1], x[2], x[3])
      }
    )
    
    return(reclassified_raster)
  }

```

STACKS ONLY RGB RASTERS
```{r}

stacking_selected_bands <- function(raster) {
  #range<-list([10:25],[30:69], [75:92],[180:204])
  red_stack <- raster[[180:204]]
  orange_stack <- raster[[75:92]]
  blue_green_stack <- raster[[30:69]]
  blue_stack <- raster[[10:25]]
  subset_raster <- stack(red_stack, orange_stack, blue_stack)
  return(subset_raster)
}

```

APPLIES BINARY MASK ONTO BAND-SUBSETTED RASTER

```{r}

applying_mask_on_subset <- function(mask, raster) {
  masked <- mask * raster
  return(masked)
}
```


CONVERTS HYPERCUBE TO 2D MATRIX
```{r}

convert_to_2d <- function(hypercube) {
  matrix2D <-
    matrix(
      hypercube,
      nrow = dim(hypercube)[1] * dim(hypercube)[2],
      ncol = dim(hypercube)[3]
    )
  return(matrix2D)
}
```

COMPLETES PCA, FINDS EIGENVALUES, USES KAISERS RULE TO DETERMINE NUMBER OF COMPONENTS
```{r}

complete_pca <- function(matrix) {
  pca_result <- prcomp(matrix, scale. = TRUE)
  
  # Extract eigenvalues
  eigenvalues <- pca_result$sdev ^ 2
  
  # Create a scree plot to visualize eigenvalues and determine the number of components
  plot(
    eigenvalues,
    type = "b",
    xlab = "Component",
    ylab = "Eigenvalue",
    main = "Scree Plot"
  )
  abline(h = 1, col = "red", lty = 2)  # Kaiser's rule threshold
  
  # Uses Kaiser-Guttman rule which suggest that "the elbow" in scree plots occurs
  # when the eigenvalues dip below 1.
  # https://datasciencewiki.net/kaisers-rule/
  num_components <- sum(eigenvalues > 1)
  return(num_components)
}
```

CONVERT ZEROS TO NANS
1) CONVERTS ALL ZEROS IN 2D MATRIX INTO NANS
```{R}
convert_to_zero<- function(matrix2D){
  nan_matrix<-gsub(0,NA, matrix2D)
  return(nan_matrix)
}
```

CHUNKY MCR TASKS
1) CREATES INITIAL CONCENTRATION/SPECTRAL GUESSES
2) COMPLETES MCR-ALS
3) PLOTS ORIGINAL DATA VS RECONSTRUCTED SPECTRA
4) OUTPUTS NUMBER OF RELEVANT COMPONENTS FROM RECONSTRUCTED DATASET

```{r}

complete_mcr_als<-function(matrix2D, component, type){
  # Step 1: Perform MCR-ALS on control_2d
  # Initialize random matrices for MCR-ALS
  initial_concentration_profiles <-
    matrix(runif(nrow(matrix2D) * component),
           nrow = nrow(matrix2D),
           ncol = component)
  initial_spectra <- matrix(runif(component * ncol(matrix2D)),
                            nrow = ncol(matrix2D),
                            ncol = component)
  
  # Run MCR-ALS on control_2d
  mcr_result <- als(
    CList = list(initial_concentration_profiles),
    S = initial_spectra,
    PsiList = list(matrix2D),
    nonnegC = TRUE,
    nonnegS = TRUE,
    maxiter = 75,
    thresh = 1e-6
  )
  # Extract the resolved spectral profiles from mcr output
  resolved_spectra <- mcr_result$S
  resolved_concentration <- mcr_result$C[[1]]
  
  reconstructed_spectra <-
    resolved_concentration %*% t(resolved_spectra)
  #
  # Compares observed_spectra, input to the MCR-ALS algorithm [n_samples, n_bands], to reconstructed results
  # Example for plotting the original vs. reconstructed data
  n_samples <- nrow(matrix2D)
  n_bands <- ncol(matrix2D)
  title <-
    paste0(type, ": Comparison of Original and Reconstructed Spectra")
  save_name <-
    paste0(
      'C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/',
      type,
      '_mcr_als.jpg'
    )
  # Plot for the first sample
  jpeg(file = save_name)
  plot(
    1:n_bands,
    matrix2D[3,],
    type = "l",
    col = "red",
    xlab = "Wavelength/Band",
    ylab = "Intensity",
    main = title
  )
  lines(1:n_bands, reconstructed_spectra[3,], col = "blue")
  legend(
    "topright",
    legend = c("Original", "Reconstructed"),
    col = c("red", "blue")
  )
  dev.off()
  #Complete PCA to find if the number of components has changed
  # # Eigenvalue calculation (using concentration profiles as an example)
  eigenvalues <- eigen(cov(reconstructed_spectra))$values
  
  # Kaiser’s rule: Select components with eigenvalues > 1
  significant_components <- which(eigenvalues > 1)
  return(list(comp = significant_components, spectra = resolved_spectra))
}
```

GAMES-HOWELL
```{r}

complete_howells_game <- function(control, treated) {
  # Combine the data into a single data frame
  compiled_data <-
    data.frame(
      Group = rep(c("Treated", "Control1", "Control2"), each = 59),
      Value = c(treated_spectra, control_spectra[, 1], control_spectra[, 2])
    )
  
  # Perform Games-Howell test
  #https://rdocumentation.org/packages/userfriendlyscience/versions/0.7.2/topics/posthocTGH
  gh_test <-
    posthocTGH(y = compiled_data$Value,
               x = compiled_data$Group,
               method = "games-howell")
  return(gh_test)
}
```

```{r Optional chunk- plots only means}

plotGamesHowell <- function(gh_test) {
  # Extract mean and confidence intervals from the test results
  means <- tapply(compiled_data$Value, compiled_data$Group, mean)
  ci <- tapply(compiled_data$Value, compiled_data$Group, function(x) {
    c(
      mean(x) - qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x)),
      mean(x) + qt(0.975, length(x) - 1) * sd(x) / sqrt(length(x))
    )
  })
  
  # Prepare data for plotting
  plot_data <- data.frame(
    Group = names(means),
    Mean = as.numeric(means),
    CI_Lower = sapply(ci, `[`, 1),
    CI_Upper = sapply(ci, `[`, 2)
  )
  
  # Create the plot
  ggplot(plot_data, aes(x = Group, y = Mean)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), width = 0.2) +
    labs(title = "Mean Spectral Values by Group with 95% Confidence Intervals",
         x = "Group", y = "Mean Spectral Value") +
    theme_minimal()
}
```
MAIN
```{r}

control<-stack('C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands/14_1_20230724_2023_07_24_12_33_04/raw_rd_rf')
treated<-stack('C:/Users/RDCRLAAU/Desktop/Plant_as_bioindicators/MCR-ALS R Processing/Test Bands/4_1_20230724_2023_07_24_09_00_14/raw_rd_rf')

control_mask<-creating_conditions(control)
control_red_cutoff<-control_mask$red
control_ob_avg<-control_mask$ob
control_r_bg_avg<-control_mask$rbg

control_reclassified_raster <- classify_raster_based_on_conditions(red_cutoff_raster = control_red_cutoff, orange_blue_avg_raster = control_ob_avg, red_bg_avg_raster = control_r_bg_avg)


treated_mask<-creating_conditions(treated)
treated_red_cutoff<-treated_mask$red
treated_ob_avg<-treated_mask$ob
treated_r_bg_avg<-treated_mask$rbg

treated_reclassified_raster <- classify_raster_based_on_conditions(red_cutoff_raster = treated_red_cutoff, orange_blue_avg_raster = treated_ob_avg, red_bg_avg_raster = treated_r_bg_avg)


plot(treated_reclassified_raster)


control_subset<-stacking_selected_bands(control)

treated_subset<-stacking_selected_bands(treated)

masked_control_subset<-applying_mask_on_subset(control_subset, control_reclassified_raster)
masked_treated_subset<-applying_mask_on_subset(treated_subset, treated_reclassified_raster)

control2d<-convert_to_2d(masked_control_subset)
treated2d<-convert_to_2d(masked_treated_subset)

initial_comp<-complete_pca(control2d)

control_mcr<-complete_mcr_als(control2d, initial_comp,'Control')
final_comp<-control_mcr$comp
control_spectra<-control_mcr$spectra
# control spectra: [1:59, 1:2]
treated_mcr<-complete_mcr_als(treated2d, final_comp, 'Treated')
treated_spectra<-treated_mcr$spectra
treated_comp<-treated_mcr$comp
# treated spectra: [1:59,1]
print(Sys.time()-start) # Takes 33.9 min to complete 
output<-complete_howells_game(final_comp, treated_comp)
plotGamesHowell
```
